### 2017-05-21 2017mergingmerges/section/02-Processes.tex

A \emph{combinator} is a template for a particular process which parameterizes it over the particular input and output streams, as well as values of configuration parameters such as the worker function used in a @map@ process. Each process implements a logical \emph{operator} --- so we use ``operator'' when describing the values being computed, but ``process'' and ``machine'' when referring to the implementation.

@@

% AR: Clumsy, but added "at L2 executes:" to force @(push...)@ onto its own line. Otherwise it goes off the right.
When the predicate is true, the instruction at label @A2@ executes:
@(push sOut1 v A3 [ l = v, f = F ])@
which pushes the value @v@ to the output stream @sOut1@ and proceeds to the instruction at label @A3@, once the variable @l@ is set to @v@ and @f@ to @F@ (False).

@@

The definition of the @merge@ combinator, which merges two input streams, is given in Figure~\ref{fig:Process:Merge}. The combinator binds the two input streams to @sIn1@ and @sIn2@, while the output stream is @sOut2@. The two heap variables @x1@ and @x2@ are used to store the last values read from each input stream. The process starts by pulling from each of the input streams. It then compares the two pulled values, and pushes the smaller of the values to the output stream. The process then drops the stream which yielded the the smaller value, then pulls from the same stream so that it can perform the comparison again.

@@

Our fusion algorithm takes two processes and produces a new one that computes the output of both. For example, suppose we need a single process that produces the output of the first two lines of our @uniquesUnion@ example back in \S\ref{s:Introduction}. The result will be a process that computes the result of both @group@ and @merge@ as if they were executed concurrently, where the first input stream of the @merge@ process is the same as the input stream of the @group@ process. In our informal description of the fusion algorithm we instantiate the parameters of each combinator with arguments of the same names.

@@

We also associate each of the joint result states with a description of whether each source machine has already pulled a value from each of its input streams. For the @F0@ case in the top of Figure~\ref{fig:Fusion:Pulls} we have ((A0, \{sIn1 = none\}), (B0, \{sIn1 = none, sIn2 = none\})). The result state @F0@ represents a combination of the two source states @A0@ and @B0@. As both @A0@ and @B0@ are the initial states of their respective machines, those machines have not yet pulled any values from their two input streams, so both `sIn1' and `sIn2' map to `none'.

@@

Finally, note that in the result states @F0@ through @F3@ the state of the input streams transitions from `none', to `pending' then to `have'. The `none' state means that we have not yet pulled a value from the associated stream. The `pending' state means we have pulled a value into the stream buffer variable (@b1@ in this case). The `have' state means that we have copied the pulled value from the stream buffer variable into the local variable used by each machine.

@@

Figure~\ref{fig:Process:Fused} shows the final result of fusing @group@ and @merge@ together. There are similar rules for handling the other combinations of instructions, but we defer the details to \S\ref{s:Fusion}.
The result has two input streams, @sIn1@ and @sIn2@, and two output streams: @sOut1@ from @group@, and @sOut2@ from @merge@. 
The shared input @sIn1@ is pulled by @merge@ at two places, and since both of these need to agree with when @group@ pulls, the @group@ instructions are duplicated at labels @F3@-@F5@ and @F13@-@F15@. The first set of instructions could be simplified to a single @push@ with constant propagation, as @f@ is initially true.

To complete the implementation of our example from \S\ref{s:Introduction} we would now proceed to fuse the process from the final line (also a @group@) into this new result process.
Although the result process has a single shared heap, the heap bindings from each fused process are guaranteed not to interfere: when we instantiated combinators to create the original source processes we introduced fresh names.

@@

% -----------------------------------------------------------------------------
\subsection{Breaking It Down}
We started with a pure functional program as in \S\ref{s:Introduction}, reimagined it as a dataflow graph, then interleaved imperative code that implemented two of the operators in that dataflow graph. Importantly, we needed to \emph{break down} the definition of each operator into imperative statements so that we could interleave their execution appropriately. We do this because the standard, single-threaded evaluation semantics of functional programs does not allow us to evaluate stream programs that contain both splits and joins in a space efficient way. Returning to the definition of @uniquesUnion@ from \S\ref{s:Introduction}, we cannot simply execute the @group@ operator on its entire input @sIn1@ before executing the @merge@ operator, as that would require us to buffer all data read from @sIn1@. Instead, during fusion we have performed the job of a concurrent scheduler at compile time. The flow of control alternates between the implementation of each of the operators, but as the instructions are interleaved directly there is no overhead due to context switching --- as there would be in a standard concurrent implementation using multiple threads.

The general approach of converting a pure functional program to a dataflow graph, then interleaving imperative statements that implement each operator was also used in prior work on Flow Fusion~\cite{lippmeier2013data}. However, in contrast to Flow Fusion and similar systems, with \mbox{Machine Fusion} we do not need to organize statements into a fixed \emph{loop anatomy} --- we simply merge them as they are. This allows us to implement a wider range of processes, including ones with nested loops that work on segmented streams.

Note that relying on lazy evaluation for an example such as @uniquesUnion@ does not solve the buffering problem. Suppose we converted each of the streams to lazy lists, and used definitions of @group@ and @merge@ that worked over these lists. As @uniquesUnion@ returns a pair of results, there is nothing preventing a consumer from demanding the first list (@sUnique@) in its entirety before demanding any of the elements from the second list (@sUnion@). If this were to happen then the runtime implementation would need to retain all elements of @sIn1@ before demanding any of @sIn2@, causing a space leak. Lazy evaluation is \emph{pull only} meaning that evaluation is driven by the consumer. The space efficiency of our fused program relies critically on the fact that operators can also \emph{push} their result values directly to their consumers, and that the consumers cannot defer the handling of these values.

>>>>

A \emph{combinator} is a template for a process which parameterizes it over the particular input and output streams, as well as values of configuration parameters such as the worker function used in a @map@ process. Each process implements a logical \emph{operator} --- so we use ``operator'' when describing the values being computed, but ``process'' and ``machine'' when referring to the implementation.

@@

% AR: Clumsy, but added "at L2 executes:" to force @(push...)@ onto its own line. Otherwise it goes off the right.
When the predicate is true, the instruction at label @A2@ executes
@(push sOut1 v A3 [ l = v, f = F ])@
which pushes the value @v@ to the output stream @sOut1@ and proceeds to the instruction at label @A3@, once the variable @l@ is set to @v@ and @f@ to @F@ (False).

@@

The definition of the @merge@ combinator, which merges two input streams, is given in Figure~\ref{fig:Process:Merge}. The combinator binds the two input streams to @sIn1@ and @sIn2@, while the output stream is @sOut2@. The two heap variables @x1@ and @x2@ store the last values read from each input stream. The process starts by pulling from each of the input streams. It then compares the two pulled values, and pushes the smaller of the values to the output stream. The process then drops the stream which yielded the the smaller value, then pulls from the same stream so that it can perform the comparison again.

@@

Our fusion algorithm takes two processes and produces a new one that computes the output of both. For example, suppose we need a single process that produces the output of the first two lines of our @uniquesUnion@ example back in \S\ref{s:Introduction}. The result will be a process that computes the result of both @group@ and @merge@ as if they were executed concurrently, where the first input stream of the @merge@ process is the same as the input stream of the @group@ process. In our informal description of the fusion algorithm we will instantiate the parameters of each combinator with arguments of the same names.

@@

We also associate each of the joint result states with a description of whether each source machine has already pulled a value from each of its input streams. For the @F0@ case at the top of Figure~\ref{fig:Fusion:Pulls} we have ((A0, \{sIn1 = none\}), (B0, \{sIn1 = none, sIn2 = none\})). The result state @F0@ represents a combination of the two source states @A0@ and @B0@. As both @A0@ and @B0@ are the initial states of their respective machines, those machines have not yet pulled any values from their two input streams, so both `sIn1' and `sIn2' map to `none'.

@@

Finally, note that in the result states @F0@ through @F3@, the state of the input streams transitions from `none', to `pending' then to `have'. The `none' state means that we have not yet pulled a value from the associated stream. The `pending' state means we have pulled a value into the stream buffer variable (@b1@ in this case). The `have' state means that we have copied the pulled value from the stream buffer variable into the local variable used by each machine.

@@

Figure~\ref{fig:Process:Fused} shows the final result of fusing @group@ and @merge@ together. There are similar rules for handling the other combinations of instructions, but we defer the details to \S\ref{s:Fusion}.
The result process has two input streams, @sIn1@ and @sIn2@, and two output streams: @sOut1@ from @group@, and @sOut2@ from @merge@. 
The shared input @sIn1@ is pulled by @merge@ instructions at two places, and since both of these need to agree with when @group@ pulls, the @group@ instructions are duplicated at @F3@-@F5@ and @F13@-@F15@. The first set of instructions could be simplified by constant propagation to a single @push@, as @f@ is initially true.

To complete the implementation of our example from \S\ref{s:Introduction} we would now fuse this result process with a process from the final line of the example (also a @group@).
Note that although the result process has a single shared heap, the heap bindings from each fused process are guaranteed not to interfere, as when we instantiate combinators to create source processes we introduce fresh names.

@@

% -----------------------------------------------------------------------------
\subsection{Breaking It Down}
We started with a pure functional program in \S\ref{s:Introduction}, reimagined it as a dataflow graph, then interleaved imperative code that implemented two of the operators in that dataflow graph. We needed to \emph{break down} the definition of each operator into imperative statements so that we could interleave their execution appropriately. We do this because the standard, single-threaded evaluation semantics of functional programs does not allow us to evaluate stream programs that contain both splits and joins in a space efficient way. Returning to the definition of @uniquesUnion@ from \S\ref{s:Introduction}, we cannot simply execute the @group@ operator on its entire input @sIn1@ before executing the @merge@ operator, as that would require us to buffer all data read from @sIn1@. Instead, during fusion we perform the job of a concurrent scheduler at compile time. In the result process the flow of control alternates between the instructions for both the @group@ and @merge@ operators, but as the instructions are interleaved directly there is no overhead due to context switching --- as there would be in a standard concurrent implementation using multiple threads.

The general approach of converting a pure functional program to a dataflow graph, then interleaving imperative statements that implement each operator was also used in prior work on Flow Fusion~\cite{lippmeier2013data}. However, in contrast to Flow Fusion and similar systems, with \mbox{Machine Fusion} we do not need to organize statements into a fixed \emph{loop anatomy} --- we simply merge them as they are. This allows us to implement a wider range of operators, including ones with nested loops that work on segmented streams.

Note that relying on lazy evaluation for @uniquesUnion@ does not eliminate the need for unbounded buffering. Suppose we converted each of the streams to lazy lists, and used definitions of @group@ and @merge@ that worked over these lists. As @uniquesUnion@ returns a pair of results, there is nothing preventing a consumer from demanding the first list (@sUnique@) in its entirety before demanding any of the elements from the second list (@sUnion@). If this were to happen then the runtime implementation would need to retain all elements of @sIn1@ before demanding any of @sIn2@, causing a space leak. Lazy evaluation is \emph{pull only} meaning that evaluation is driven by the consumer. The space efficiency of our fused program relies critically on the fact that processes can also \emph{push} their result values directly to their consumers, and that the consumers cannot defer the handling of these values.

### 2017-05-21 2017mergingmerges/section/A1-Finite.tex

\begin{figure}
\begin{tabbing}
MMMMMM \TABDEF @MMMMM@  \TABSKIP $\Chan$ \TABSKIP $\Chan$ \TABSKIP $\Next$ \TABSKIP \kill
\Instr
    \> ::=\> @push@  \> \Chan  \> \Exp  \> \Next \\
    \TABALT  @drop@  \> \Chan  \>       \> \Next \\
    \TABALT  @case@  \> \Exp   \> \Next \> \Next \\
    \TABALT  @jump@  \>        \>       \> \Next \\
    \\
    \TABALT  @pull@  \> \Chan  \> \Var  \> \Next \> \Next \\
    \TABALT  @close@ \> \Chan  \>       \> \Next \\
    \TABALT  @exit@ 
\\
\\
$\InputState_F$ \> ::=  \> ~~~ $@none@_F ~|~ @pending@_F ~|~ @have@_F ~|~ @closed@_F$
\end{tabbing}
\caption{Finite instructions}
\label{fig:Finite:Instr}
\end{figure}

@@

Also in Figure~\ref{fig:Finite:Instr}, the static input state used for fusion ($\InputState_F$) must now track closed streams. The new constructor $@closed@_F$ denotes that the stream is closed, while the rest is unchanged.

@@

Figure~\ref{fig:Finite:tryStepPair} shows the modified $\ti{tryStepPair}$ function. This function uses the same heuristics to decide which process to execute when both can progress, but now that the processes can finish with @exit@, we must take care to only finish the fused process once \emph{both} source processes are finished. The (DeferExit1) and (DeferExit2) clauses achieve this by forcing the other process to run if one is an @exit@. Once both processes are finished, both new clauses will fail while (DeferPull1) succeeds, using the @exit@ from the first process. Another way to think of this is that if either process has work to do, the fused process still has work to do.

\begin{figure}
\begin{tabbing}
M \= M \= M \= MMMMMMMMMMMMMMMMMMM \= \kill
$\ti{tryStepPair} ~:~ \ChanTypeMap$ \\
\> \> $\to \Label_1 \to \Instr$ ~~~
            $\to \Label_1 \to \Instr$ \\
\> \> $\to \Maybe~\Instr$ \\
$\ti{tryStepPair} ~\cs~l_p~i_p~l_q~i_q$ \\

\> $~|~i_p'~\in~\ti{tryStep}~\cs~l_p~i_p~l_q ~\wedge~ i_q'~\in~\ti{tryStep}~\cs~l_q~i_q~l_p$ \\
\> ~~ $\wedge~@exit@~\not\in~i_p'$ 
   ~~~~~ $\to~i_p'$ 
\> \> \> \note{DeferExit1} 
\\[0.5ex]

\> $~|~i_p'~\in~\ti{tryStep}~\cs~l_p~i_p~l_q ~\wedge~i_q'~\in~\ti{tryStep}~\cs~l_q~i_q~l_p$ \\
\> ~~ $\wedge~@exit@~\not\in~i_q'$ 
   ~~~~~ $\to~\ti{swaplabels}~i_q'$ 
\> \> \> \note{DeferExit2} 
\\[0.5ex]

\> $~|~i_p'~\in~\ti{tryStep}~\cs~l_p~i_p~l_q ~\wedge~@jump@~(l,u)~\in~i_p'$ \\
\> $\to~i_p'$
\> \> \> \note{PreferJump1} 
\\[0.5ex]

\> $~|~i_q'~\in~\ti{tryStep}~\cs~l_q~i_q~l_p ~\wedge~@jump@~(l,u)~\in~i_q'$ \\
\> $\to~\ti{swaplabels}~i_q'$ 
\> \> \> \note{PreferJump2}
\\[0.5ex]

\> $~|~i_p'~\in~\ti{tryStep}~\cs~l_p~i_p~l_q ~\wedge~ i_q'~\in~\ti{tryStep}~\cs~l_q~i_q~l_p$ \\
\> ~~ $\wedge~@pull@~c~x~(l,u)~\not\in~i_p'$ 
   ~~~~~ $\to~i_p'$ 
\> \> \> \note{DeferPull1} 
\\[0.5ex]

\> $~|~i_p'~\in~\ti{tryStep}~\cs~l_p~i_p~l_q ~\wedge~i_q'~\in~\ti{tryStep}~\cs~l_q~i_q~l_p$ \\
\> ~~ $\wedge~@pull@~c~x~(l,u)~\not\in~i_q'$ 
   ~~~~~ $\to~\ti{swaplabels}~i_q'$ 
\> \> \> \note{DeferPull2} 
\\[0.5ex]

\> $~|~i_p'~\in~\ti{tryStep}~\cs~l_p~i_p~l_q$ ~~~~~ $\to~i_p'$ 
\> \> \> \note{Run1} 
\\[0.5ex]

\> $~|~i_q'~\in~\ti{tryStep}~\cs~l_q~i_q~l_p$ ~~~~~ $\to~\ti{swaplabels}~i_q'$
\> \> \> \note{Run2} \\

\end{tabbing}
\caption{Fusion step coordination for a pair of processes.}
\label{fig:Finite:tryStepPair}
\end{figure}

@@

\begin{figure*}
\begin{tabbing}
M \= M \= MMMMMMMMMMMMMMMMMMMMMM \= MMMMMMMMMMMMMMMMMMMMMMMMMMMMMM \= \kill
$\ti{tryStep} ~:~ \ChanTypeMap \to \Label_1 \to \Instr \to \Label_1 \to \Maybe~\Instr$ \\
$\ti{tryStep} ~\cs~(l_p,s_p)~i_p~(l_q,s_q)~=~@match@~i_p~@with@$ \\

\> $@jump@~(l',u')$ 
\> \> $\to~@jump@~
      \nextStep
        {l'}{s_p}
        {l_q}{s_q}
        {u'}
      $ 
\> \note{LocalJump}
\\[1ex]

\> $@case@~e~(l'_t,u'_t)~(l'_f,u'_f)$
\> \> $\to~@case@~e~
      \nextStep
        {l'_t}{s_p}
        {l_q}{s_q}
        {u'_t}
      ~
      \nextStep
        {l'_f}{s_p}
        {l_q}{s_q}
        {u'_f}
      $ 
\> \note{LocalCase}
\\[1ex]

\> $@push@~c~e~(l',u')$ \\
\> \> $~|~\cs[c]=@out1@$ 
\> $\to~@push@~c~e~
      \nextStep
        {l'}
          {s_p}
        {l_q}
          {s_q}
        {u'}
      $ 
\> \note{LocalPush}\\

\> \> $~|~\cs[c]=@in1out1@ ~\wedge~ s_q[c]=@none@_F$ 
\> $\to~@push@~c~e~
      \nextStep
        {l'}
          {s_p}
        {l_q}
          {\HeapUpdateOne{c}{@pending@_F}{s_q}}
        {\HeapUpdateOne{@chan@~c}{e}{u'}}
      $
\> \note{SharedPush}
\\[1ex]

\> $@drop@~c~(l',u')$ \\
\> \> $~|~\cs[c]=@in1@$
\> \hspace{2em} $\to~@drop@~c~
      \nextStep
        {l'}
          {s_p}
        {l_q}
          {s_q}
        {u'}
      $
\> \note{LocalDrop} \\

\> \> $~|~\cs[c]=@in1out1@$
\> \hspace{2em} $\to~@jump@~
      \nextStep
        {l'}
          {\HeapUpdateOne{c}{@none@_F}{s_p}}
        {l_q}
          {s_q}
        {u'}
      $
\> \note{ConnectedDrop}\\

\> \> $~|~\cs[c]=@in2@ ~\wedge~ (s_q[c]=@have@_F \vee s_q[c]=@pending@_F)$ 
\> \hspace{2em} $\to~@jump@~
      \nextStep
        {l'}
          {\HeapUpdateOne{c}{@none@_F}{s_p}}
        {l_q}
          {s_q}
        {u'}
      $
\> \note{SharedDropOne}\\

\> \> $~|~\cs[c]=@in2@ ~\wedge~ s_q[c]=@none@_F$
\> \hspace{2em} $\to~@drop@~c~
      \nextStep
        {l'}
          {\HeapUpdateOne{c}{@none@_F}{s_p}}
        {l_q}
          {s_q}
        {u'}
      $
\> \note{SharedDropBoth}
\\[1ex]

@@

\> $@pull@~c~x~(l'_o,u'_o)~(l'_c,u'_c)$ \\
\> \> $~|~\cs[c]=@in1@$ 
\> $\to~@pull@~c~x~
      \nextStep
        {l'_o}{s_p}
        {l_q}{s_q}
        {u'_o}
      ~
      \nextStep
        {l'_c}{s_p}
        {l_q}{s_q}
        {u'_c}
    $ 
\> \note{LocalPull}
\\[1ex]

\> \> $~|~(\cs[c]=@in2@ \vee \cs[c]=@in1out1@) ~\wedge~ s_p[c]=@pending@_F$ \\
\> \> $\to~@jump@~
      \nextStep
        {l'_o}
          {\HeapUpdateOne{c}{@have@_F}{s_p}}
        {l_q}
          {s_q}
        {\HeapUpdateOne{x}{@chan@~c}{u'_o}}
        $ 
\> \> \note{SharedPull} 
\\[1ex]

\> \> $~|~\cs[c]=@in2@ ~\wedge~ s_p[c]=@none@_F ~\wedge~ s_q[c]=@none@_F$ \\
\> \> $\to~@pull@~c~(@chan@~c)~
      \nextStep
        {l_p}
          {\HeapUpdateOne{c}{@pending@_F}{s_p}}
        {l_q}
          {\HeapUpdateOne{c}{@pending@_F}{s_q}}
        {[]}$
      \\
\> \> @                @
      $\nextStep
        {l_p}
          {\HeapUpdateOne{c}{@closed@_F}{s_p}}
        {l_q}
          {\HeapUpdateOne{c}{@closed@_F}{s_q}}
        {[]}
  $
\> \> \note{SharedPullInject}
\\[1ex]

\> \> $~|~(\cs[c]=@in2@ \vee \cs[c]=@in1out1@) ~\wedge~ s_p[c]=@closed@_F$ \\
\> \> $\to~@jump@~
      \nextStep
        {l'_c}{s_p}
        {l_q}{s_q}
        {u'_c}
  $
\> \> \note{SharedPullClosed}
\\[1ex]

\> $@close@~c~(l',u')$ \\
\> \> $~|~\cs[c]=@out1@$ 
\> $\to~@close@~c~
      \nextStep
        {l'}{s_p}
        {l_q}{s_q}
        {u'}
    $ 
\> \note{LocalClose}
\\

\> \> $~|~\cs[c]=@in1out1@ ~\wedge~ s_q[c]=@none@_F$ 
\> $\to~@close@~c~
      \nextStep
        {l'}{s_p}
        {l_q}{\HeapUpdateOne{c}{@closed@_F}{s_q}}
        {u'}
    $ 
\> \note{SharedClose}
\\[1ex]

\> $@exit@$
\> 
\> $\to~@exit@$
\> \note{LocalExit}
\\

>>>>

\begin{figure}
\begin{tabbing}
MMMMMM \TABDEF @MMMMM@  \TABSKIP $\Chan$ \TABSKIP $\Chan$ \TABSKIP $\Next$ \TABSKIP \kill
\Instr
    \> ::=\> @push@  \> \Chan  \> \Exp  \> \Next \\
    \TABALT  @drop@  \> \Chan  \>       \> \Next \\
    \TABALT  @case@  \> \Exp   \> \Next \> \Next \\
    \TABALT  @jump@  \>        \>       \> \Next \\
    \\
    \TABALT  @pull@  \> \Chan  \> \Var  \> \Next \> \Next \\
    \TABALT  @close@ \> \Chan  \>       \> \Next \\
    \TABALT  @exit@ 
\\
\\
$\InputStateF$ \> ::=  \> ~~~ $@none@_F ~|~ @pending@_F ~|~ @have@_F ~|~ @closed@_F$
\end{tabbing}
\caption{Finite instructions}
\label{fig:Finite:Instr}
\end{figure}

@@

Also in Figure~\ref{fig:Finite:Instr}, the static input state used for fusion ($\InputStateF$) must now track closed streams. The new constructor $@closed@_F$ denotes that the stream is closed, while the rest is unchanged.

@@

Figure~\ref{fig:Finite:tryStepPair} shows the modified $\ti{tryStepPair}$ function. This function uses the same heuristics to decide which process to execute when both can progress, but now that the processes can finish with @exit@, we must take care to only finish the fused process once \emph{both} source processes are finished. The (DeferExit1) and (DeferExit2) clauses achieve this by forcing the other process to run if one is an @exit@. Once both processes are finished, both new clauses will fail while (Run1) succeeds, using the @exit@ from the first process. Another way to think of this is that if either process has work to do, the fused process still has work to do.

\begin{figure}
\begin{tabbing}
\ti{tryStepPair}\=$~\to~$\=M\kill

$\ti{tryStepPair} ~:~ \ChanTypeMap$ \\
\> $\to$ \> $\LabelF \to \Instr \to \LabelF \to \Instr$ \\
\> $\to$ \> $\Maybe~\Instr$ \\

M \= $~|~(@pull@~\_~\_~\_ \gets i_p')~$ \= $\to$ \= $\Just (\ti{swaplabels}~i_q')~$ \= M \=\kill
$\ti{tryStepPair} ~\cs~l_p~i_p~l_q~i_q~=$ \\
\> $@match@~ (\ti{tryStep}~\cs~l_p~i_p~l_q,~\ti{tryStep}~\cs~l_q~i_q~l_p) ~@with@$ \\
\> $(\Just i_p',~\Just i_q')$ \\

\> @ @$|~@exit@ \gets i_q'$ \> $\to$ \> $\Just i_p'$
\> \note{DeferExit1} \\[0.5ex]

\> @ @$|~@exit@ \gets i_p'$ \> $\to$ \> $\Just (\ti{swaplabels}~i_q')$
\> \note{DeferExit2} \\[0.5ex]

\> @ @$|~@jump@~\_ \gets i_p'$ \> $\to$ \> $\Just i_p'$
\> \note{PreferJump1} \\[0.5ex]

\> @ @$|~@jump@~\_ \gets i_q'$ \> $\to$ \> $\Just (\ti{swaplabels}~i_q')$
\> \note{PreferJump2} \\[0.5ex]

\> @ @$|~@pull@~\_~\_~\_ \gets i_q'$ \> $\to$ \> $\Just i_p'$
\> \note{DeferPull1} \\[0.5ex]

\> @ @$|~@pull@~\_~\_~\_ \gets i_p'$ \> $\to$ \> $\Just (\ti{swaplabels}~i_q')$
\> \note{DeferPull2} \\[0.5ex]

\> $(\Just i_p',~\_)$ \> $\to$ \> $\Just i_p'$
\> \note{Run1} \\[0.5ex]

\> $(\_, ~\Just i_q')$ \> $\to$ \> $\Just (\ti{swaplabels}~i_q')$
\> \note{Run2} \\[0.5ex]

\> $(\Nothing, ~\Nothing)$ \> $\to$ \> $\Nothing$
\> \note{Deadlock}
\end{tabbing}
\caption{Fusion step coordination for a pair of processes.}
\label{fig:Finite:tryStepPair}
\end{figure}

@@

\begin{figure*}
\begin{tabbing}
M \= M \= MMMMMMMMMMMMMMMMMMMMMM \= MMMMMMMMMMMMMMMMMMMMMMMMMMMMMM \= \kill
$\ti{tryStep} ~:~ \ChanTypeMap \to \LabelF \to \Instr \to \LabelF \to \Maybe~\Instr$ \\
$\ti{tryStep} ~\cs~(l_p,s_p)~i_p~(l_q,s_q)~=~@match@~i_p~@with@$ \\

\> $@jump@~(l',u')$ 
\> \> $\to~\Just (@jump@~
      \nextStep
        {l'}{s_p}
        {l_q}{s_q}
        {u'})
      $ 
\> \note{LocalJump}
\\[1ex]

\> $@case@~e~(l'_t,u'_t)~(l'_f,u'_f)$
\> \> $\to~\Just (@case@~e~
      \nextStep
        {l'_t}{s_p}
        {l_q}{s_q}
        {u'_t}
      ~
      \nextStep
        {l'_f}{s_p}
        {l_q}{s_q}
        {u'_f})
      $ 
\> \note{LocalCase}
\\[1ex]

\> $@push@~c~e~(l',u')$ \\
\> \> $~|~\cs[c]=@out1@$ 
\\
\> \> $\to~\Just (@push@~c~e~
      \nextStep
        {l'}
          {s_p}
        {l_q}
          {s_q}
        {u'})
      $ 
\> \> \note{LocalPush}\\

\> \> $~|~\cs[c]=@in1out1@ ~\wedge~ s_q[c]=@none@_F$ 
\\
\> \> $\to~\Just (@push@~c~e~
      \nextStep
        {l'}
          {s_p}
        {l_q}
          {\HeapUpdateOne{c}{@pending@_F}{s_q}}
        {\HeapUpdateOne{@chan@~c}{e}{u'}})
      $
\> \> \note{SharedPush}
\\[1ex]

\> $@drop@~c~(l',u')$ \\
\> \> $~|~\cs[c]=@in1@$
\> \hspace{2em} $\to~\Just (@drop@~c~
      \nextStep
        {l'}
          {s_p}
        {l_q}
          {s_q}
        {u'})
      $
\> \note{LocalDrop} \\

\> \> $~|~\cs[c]=@in1out1@$
\> \hspace{2em} $\to~\Just (@jump@~
      \nextStep
        {l'}
          {\HeapUpdateOne{c}{@none@_F}{s_p}}
        {l_q}
          {s_q}
        {u'})
      $
\> \note{ConnectedDrop}\\

\> \> $~|~\cs[c]=@in2@ ~\wedge~ (s_q[c]=@have@_F \vee s_q[c]=@pending@_F)$ 
\> \hspace{2em} $\to~\Just (@jump@~
      \nextStep
        {l'}
          {\HeapUpdateOne{c}{@none@_F}{s_p}}
        {l_q}
          {s_q}
        {u'})
      $
\> \note{SharedDropOne}\\

\> \> $~|~\cs[c]=@in2@ ~\wedge~ s_q[c]=@none@_F$
\> \hspace{2em} $\to~\Just (@drop@~c~
      \nextStep
        {l'}
          {\HeapUpdateOne{c}{@none@_F}{s_p}}
        {l_q}
          {s_q}
        {u'})
      $
\> \note{SharedDropBoth}
\\[1ex]

@@

\> $@pull@~c~x~(l'_o,u'_o)~(l'_c,u'_c)$ \\
\> \> $~|~\cs[c]=@in1@$ 
\\
\> \> $\to~\Just (@pull@~c~x~
      \nextStep
        {l'_o}{s_p}
        {l_q}{s_q}
        {u'_o}
      ~
      \nextStep
        {l'_c}{s_p}
        {l_q}{s_q}
        {u'_c})
    $ 
\> \> \note{LocalPull}
\\[1ex]

\> \> $~|~(\cs[c]=@in2@ \vee \cs[c]=@in1out1@) ~\wedge~ s_p[c]=@pending@_F$ \\
\> \> $\to~\Just (@jump@~
      \nextStep
        {l'_o}
          {\HeapUpdateOne{c}{@have@_F}{s_p}}
        {l_q}
          {s_q}
        {\HeapUpdateOne{x}{@chan@~c}{u'_o}})
        $ 
\> \> \note{SharedPull} 
\\[1ex]

\> \> $~|~\cs[c]=@in2@ ~\wedge~ s_p[c]=@none@_F ~\wedge~ s_q[c]=@none@_F$ \\
\> \> $\to~\Just (@pull@~c~(@chan@~c)~
      \nextStep
        {l_p}
          {\HeapUpdateOne{c}{@pending@_F}{s_p}}
        {l_q}
          {\HeapUpdateOne{c}{@pending@_F}{s_q}}
        {[]}$
      \\
\> \> @                     @
      $\nextStep
        {l_p}
          {\HeapUpdateOne{c}{@closed@_F}{s_p}}
        {l_q}
          {\HeapUpdateOne{c}{@closed@_F}{s_q}}
        {[]})
  $
\> \> \note{SharedPullInject}
\\[1ex]

\> \> $~|~(\cs[c]=@in2@ \vee \cs[c]=@in1out1@) ~\wedge~ s_p[c]=@closed@_F$ \\
\> \> $\to~\Just (@jump@~
      \nextStep
        {l'_c}{s_p}
        {l_q}{s_q}
        {u'_c})
  $
\> \> \note{SharedPullClosed}
\\[1ex]

\> $@close@~c~(l',u')$ \\
\> \> $~|~\cs[c]=@out1@$ 
\> $\to~\Just (@close@~c~
      \nextStep
        {l'}{s_p}
        {l_q}{s_q}
        {u'})
    $ 
\> \note{LocalClose}
\\

\> \> $~|~\cs[c]=@in1out1@ ~\wedge~ s_q[c]=@none@_F$ 
\> $\to~\Just (@close@~c~
      \nextStep
        {l'}{s_p}
        {l_q}{\HeapUpdateOne{c}{@closed@_F}{s_q}}
        {u'})
    $ 
\> \note{SharedClose}
\\[1ex]

\> $@exit@$
\> 
\> $\to~\Just @exit@$
\> \note{LocalExit}
\\[1ex]

% \> @otherwise@ \>
\> $\_$ \> $~|~ @otherwise@ $
\> $\to ~ \Nothing$
\> \note{Blocked}
